\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[margin=1in]{geometry}
\usepackage{setspace}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{xcolor}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=blue,citecolor=blue]{hyperref}

\setstretch{1.08}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.6em}
\setlist[itemize]{leftmargin=1.6em}
\setlist[enumerate]{leftmargin=1.8em}

\newcommand{\journalname}{Virtual Reality \& Intelligent Hardware}
\newcommand{\manuscriptnumber}{VRIH-D-26-00032}
\newcommand{\manuscripttitle}{MeDeA: Decomposable Attention for Intrinsically Interpretable ECG Diagnosis}
\newcommand{\commentheader}{\textbf{Comment.}}
\newcommand{\responseheader}{\textbf{Response.}}
\newcommand{\revisionheader}{\textbf{Revision in the manuscript.}}

\begin{document}

\begin{center}
{\Large \textbf{Response to the Editor and Reviewers}}\\[0.6em]
{\normalsize \textbf{Journal:} \journalname}\\
{\normalsize \textbf{Manuscript No.:} \manuscriptnumber}\\
{\normalsize \textbf{Title:} \manuscripttitle}\\
{\normalsize \textbf{Decision:} Major Revision}
\end{center}

We sincerely thank the Editor and the Reviewers for their careful assessment of our manuscript and for the constructive suggestions. We have revised both the manuscript text and the presentation of the experimental evidence accordingly. In particular, we have clarified the relation between the backbone and the attention module, made the evaluation protocol more explicit, corrected the citation mismatch, strengthened the discussion on generalization and clinical limitations, and moderated claims that were stronger than the currently documented evidence supports. The revised manuscript is therefore more precise in both methodological description and empirical interpretation.

\section*{Summary of Major Revisions}

\begin{itemize}
    \item We rewrote the method description to make the data flow explicit, including the tensor dimensions from the ECG input to the backbone output, the query/key/value projections, the query-wise attention maps, and the final class-level explanation maps.
    \item We clarified that the multi-query vectors are global learnable parameters rather than input-derived features, and we removed the ambiguity around the temperature notation by aligning the manuscript with the implemented scaled dot-product form.
    \item We expanded the experimental protocol description to state the dataset split organization, the default optimization settings, the lack of explicit data augmentation in the released scripts, the seed control, and the boundary between in-house runs and literature-reported comparison methods.
    \item We added an efficiency-oriented comparison based on reproducible quantities available from the current codebase, including trainable parameter count, FP32 weight footprint, and a lightweight forward-latency benchmark, and we toned down the earlier deployment claims accordingly.
    \item We revised the narrative of the comparative experiments to avoid implying universal superiority. The revised manuscript now presents MeDeA as a competitive and intrinsically interpretable alternative rather than as an unqualified state-of-the-art method.
    \item We strengthened the discussion of limitations and clinical translation, especially regarding cross-hospital shift, device shift, sampling-rate mismatch, retrospective public data, and dependence on reliable annotation.
    \item We corrected the reference mismatch around IMLENet and rechecked the bibliography to replace arXiv-only citations with peer-reviewed versions where available.
    \item We revised the figure descriptions so that Figures 1--3 are easier to interpret and reproduce, including clearer captions, explicit shape annotations, and a precise explanation of the multi-label prediction-rate matrix.
\end{itemize}

\section*{Response to the Editor}

\commentheader\ The Editor requested that all reviewer comments be addressed point by point and that every substantive revision be documented clearly.

\responseheader\ We have carefully revised the manuscript in accordance with all comments from the two reviewers. The revised version now provides a clearer description of the MeDeA architecture, explicitly separates code-supported in-house experiments from literature-reported comparisons, corrects the citation issue noted by Reviewer \#1, strengthens the discussion of external validity and clinical limitations, and substantially improves the description of the figures. We have also moderated the performance claims so that they match the evidence documented by the current implementation and experiment logs.

\revisionheader\ The revised manuscript now includes a more explicit methodological narrative, a clarified experimental protocol subsection, an added efficiency note, improved figure captions, a corrected bibliography entry for IMLENet, and an expanded discussion/limitations section. The present response letter documents these changes point by point below.

\section*{Response to Reviewer \#1}

\subsection*{Comment 1.1: Class imbalance handling in the multi-label setting}

\commentheader\ Although BCE is reported for the multi-label task, the manuscript does not explain how class imbalance is handled, especially when class prevalence varies substantially across labels.

\responseheader\ We thank the reviewer for raising this important point. After rechecking both the manuscript and the released implementation, we found that our original wording was not sufficiently precise and could be misleading. For the main PTB-XL multi-label experiments reported in this work, the training loss is the standard \texttt{BCEWithLogitsLoss()} without \texttt{pos\_weight}. In other words, although class imbalance does exist in PTB-XL, the main experiments do not apply explicit positive-class reweighting in BCE, nor do they use focal loss, resampling, or other class-balancing strategies. By contrast, a separate single-label script uses \texttt{CrossEntropyLoss()} for a different setting.

Therefore, the original description in the Methods section suggesting that $w_c$ is a positive class weight used to handle imbalance does not accurately reflect the released implementation of the main PTB-XL experiments. We have corrected this in the revised manuscript by removing the ambiguous wording and aligning the loss description with the actual implementation. We retain the prevalence statistics in Table~1 to make the degree of class imbalance transparent, but we now state clearly that these statistics are reported to characterize the dataset rather than to indicate that explicit loss reweighting was used in the main training pipeline. We also note this design choice as a limitation, since the absence of explicit imbalance handling may affect minority-class performance.

\revisionheader\ We revised the loss-function description so that it now matches the released PTB-XL implementation, explicitly stating that the main experiments use standard \texttt{BCEWithLogitsLoss()} without \texttt{pos\_weight}. We also removed wording that could imply explicit class-weighted training, retained prevalence statistics only as a characterization of the dataset imbalance, and added a note discussing the absence of explicit imbalance handling as a limitation.

\subsection*{Comment 1.2: Relationship between the 1D ResNet backbone and the MQAH module}

\commentheader\ The relation between the backbone and MQAH, including feature-dimension matching and the effect of averaging and upsampling attention maps, is not sufficiently clear.

\responseheader\ We agree and have substantially clarified this point. The revised manuscript now explains the backbone-to-MQAH computation and the explanation pipeline as an explicit sequence of tensor transformations. Specifically, the input ECG is first processed by the shared 1D ResNet backbone to produce a compressed feature sequence $\mathbf{X}_{\mathrm{feat}} \in \mathbb{R}^{B \times C \times L'}$. For the main PTB-XL setting discussed in this work, we now provide a concrete example in the manuscript: the input has shape $(B,12,1000)$ and the backbone output is defined on a compressed temporal axis with $L'=32$. Each disease-specific MQAH head then computes attention over this feature sequence using multiple learnable queries and produces the class logit together with multiple query-wise attention maps.

We further clarify the role of averaging. The averaging over the query dimension is used only to aggregate the multiple query-specific responses within the same class head into a single class-level explanation map, thereby improving readability. It is not an additional inference module and does not change the decision pathway, although it can make the displayed explanation smoother than an individual query map because it summarizes multiple query-specific views.

We also clarify the role of upsampling. Because the attention map is defined on the downsampled temporal axis of length $L'$, we linearly interpolate it back to the original signal length $L$ only for alignment with the raw ECG waveform in visualization. In the PTB-XL example, this corresponds to mapping the explanation from length $32$ back to length $1000$. This step does not introduce new diagnostic information or increase the true temporal resolution of the explanation; the effective resolution remains limited by the backbone output length $L'$. Related clarifications have now been added to the backbone description, the MQAH subsection, the explanation-generation subsection, and the corresponding figure captions.

\revisionheader\ We rewrote the architecture description, added tensor-shape annotations and a concrete PTB-XL example to the text and to Figures 1--2, and clarified that the explanation pipeline is ``query-wise attention $\rightarrow$ query averaging $\rightarrow$ linear interpolation for visualization only,'' with the effective explanation resolution still limited by the backbone output length.

\subsection*{Comment 2.1: Fairness and reproducibility of Table 3}

\commentheader\ Table 3 lacks key details such as training settings, model size, data augmentation, and whether the same data split was adopted across methods.

\responseheader\ We agree that the original presentation was not sufficiently explicit. In the revised manuscript, we now separate the issues of protocol specification, model size, and comparison fairness more clearly. First, we added a dedicated protocol description for the main in-house PTB-XL runs produced by the released MeDeA-family scripts, including the processed PTB-XL 100 Hz data setting, random seed 42, AdamW optimizer, learning rate $10^{-3}$, weight decay $10^{-4}$, default batch size, early-stopping patience, default thresholding, and the fact that no explicit data augmentation is used in these scripts. Second, model-size information is no longer left implicit: reproducible parameter counts and FP32 weight footprints are now reported separately in the efficiency analysis subsection rather than being mixed into the protocol description.

Most importantly, we now distinguish between directly comparable in-house results and contextual reference results. The directly comparable table in the revision is restricted to runs generated under the local MeDeA-family PTB-XL pipeline, whereas repository-wide baselines or literature-reported methods are described as contextual comparisons unless the same split organization and training protocol can be verified explicitly. We therefore no longer use wording that suggests that every entry in the comparison table was obtained under a perfectly identical split and tuning setup.

\revisionheader\ We added an ``Experimental Protocol'' description for the main in-house PTB-XL MeDeA-family runs, explicitly reporting optimizer, learning rate, weight decay, batch size, stopping rule, seed control, thresholding behavior, environment note, and augmentation status. We also moved model-size information to a separate efficiency subsection and revised the comparison text so that directly comparable in-house runs are clearly separated from contextual repository-wide or literature-reported results.

\subsection*{Comment 2.2: Split consistency and tuning protocol for ECGNet, IMLENet, and other baselines}

\commentheader\ The manuscript should clarify whether the same split and a comparable hyperparameter tuning protocol were used for ECGNet, IMLENet, and the other baselines.

\responseheader\ We thank the reviewer for raising this point. We agree that, if ECGNet, IMLENet, and other external methods are to be treated as strictly comparable baselines, the ideal approach would be to rerun them under a unified split definition, threshold-selection policy, and hyperparameter protocol. We did assess this possibility during revision. However, based on the information available to us at the time of revision, these methods do not provide directly reproducible official implementations, and the published papers do not disclose enough detail to support a faithful protocol-matched rerun. Under these circumstances, reimplementing them ourselves while filling in undisclosed settings would likely introduce additional implementer bias rather than improve objectivity.

For this reason, the revised manuscript no longer presents these external results as fully protocol-matched benchmarks. Instead, we describe them as literature-reported task-level reference results. The core controlled evidence in the paper is now positioned more explicitly around the in-house model-family comparisons and ablation experiments conducted under a unified local protocol.

\revisionheader\ We revised the comparison text so that external methods such as ECGNet, IMLENet, and MSW-TN are no longer described as fully protocol-matched baselines when such equivalence cannot be verified. The revised manuscript now treats them as literature-reported task-level references, while positioning the unified in-house model-family comparisons and ablation experiments as the main controlled evidence.

\subsection*{Comment 2.3: Missing efficiency analysis}

\commentheader\ The manuscript argues for trustworthiness and computational feasibility, but it does not report parameter count, FLOPs, latency, or memory consumption.

\responseheader\ We agree that the original claim was under-supported. In the revised manuscript, we therefore add a compact efficiency note based only on quantities that can be measured reproducibly from the current repository snapshot: trainable parameter count, FP32 weight footprint, and a lightweight batch-1 CPU forward-latency benchmark for a $(12 \times 1000)$ input. These quantities do not replace a full systems study, but they do provide a controlled and reproducible basis for discussing computational cost across representative local variants and a Transformer baseline.

This additional analysis allows us to narrow our efficiency claim in a more evidence-based way. The results show that MeDeA is not uniformly the lightest model, especially relative to smaller local variants, but it remains far more practical than a plain Transformer baseline in forward latency while providing intrinsically decomposed class-wise attention. At the same time, we now state explicitly that the current revision does not provide a complete FLOPs or peak-memory benchmark, so these numbers should be interpreted as efficiency-oriented reference measurements rather than as an exhaustive deployment study.

\begin{center}
\small
\begin{tabular}{lrrr}
\toprule
Model & Parameters & FP32 Weight Footprint & Forward Latency (ms) \\
\midrule
MeDeA & 5,986,442 & 22.84 MB & 12.11 \\
MeDeA-Single-Query & 411,205 & 1.57 MB & 3.22 \\
Inception Baseline & 672,965 & 2.57 MB & 9.30 \\
Transformer Baseline & 1,202,053 & 4.59 MB & 157.73 \\
\bottomrule
\end{tabular}
\end{center}

We emphasize in the revised text that these numbers correspond to reproducible weight size and forward-pass behavior; they should not be over-interpreted as a full systems benchmark. We therefore removed any wording that previously implied an exhaustive FLOPs/peak-memory study when such a study was not actually documented in the original manuscript.

\revisionheader\ We added the efficiency comparison above to the revised manuscript, explicitly reporting parameter count, FP32 weight footprint, and lightweight forward latency as reproducible measurements from the current codebase. We also removed any wording that could imply a complete FLOPs or peak-memory study and softened the discussion of ``computational feasibility'' so that it matches the available evidence.

\subsection*{Comment 3.1: Clinical applicability and generalization}

\commentheader\ The evaluation is mainly based on public retrospective datasets; the manuscript should discuss robustness across hospitals, devices, and sampling rates, as well as dependence on annotation quality.

\responseheader\ We fully agree. The revised manuscript now states explicitly that the present evidence is limited to retrospective public datasets and is therefore insufficient to establish clinical readiness or deployment robustness. We added a dedicated discussion of external-validity limitations covering at least four points: domain shift across hospitals, acquisition differences across devices and lead setups, sampling-rate mismatch, and sensitivity to label quality or annotation noise. We also make clear that future validation must go beyond the current PTB-XL setting and include prospective, cross-site, and cross-device evaluation before stronger clinical claims can be made.

In addition, we clarify the role of interpretability more carefully. The intrinsic explanation mechanism should be understood as a model-internal explanatory aid that may improve transparency, but it is not a substitute for external clinical validation and does not by itself guarantee robustness under real-world distribution shift.

\revisionheader\ We expanded the Discussion/Limitations section and revised the conclusion so that the manuscript now explicitly limits the current evidence to retrospective public datasets, discusses hospital/device/sampling-rate/annotation-quality risks, and states that prospective cross-site validation is still required before making stronger clinical-applicability claims.

\subsection*{Comment 4.1: Citation mismatch around IMLENet}

\commentheader\ IMLENet [27] appears to be mismatched with a non-ECG paper, suggesting a citation or numbering error.

\responseheader\ Thank you for pointing out this error. We corrected the mismatched IMLENet citation in the revised manuscript and rechecked the ECG-related comparison references for entry and numbering consistency. Where a peer-reviewed version could be verified, we also preferred that version over an unnecessary arXiv-only citation.

\revisionheader\ The revised manuscript now records the corrected IMLENet citation and the accompanying bibliography cleanup so that the cited method matches the intended ECG comparison more consistently.

\subsection*{Comment 5.1: Figure 1 and Figure 2 captions need more detail}

\commentheader\ The captions and descriptions of Figures 1 and 2 should indicate feature-map dimensions or attention output sizes to improve reproducibility.

\responseheader\ We agree and revised both the figure captions and the corresponding main-text description. In the revised manuscript, Figure 1 now makes the implemented tensor interface explicit by identifying the PTB-XL input shape $(B,12,1000)$, the shared backbone output $(B,512,32)$, and the disease-specific outputs. Figure 2 likewise now states the learnable query shape $(H,d)$, the feature-derived key/value shape $(B,L',d)$, the query-wise attention-score shape $(B,H,L')$, and the final query-averaged class-level explanation shape $(B,L')$. This change is intended to make the figures reproducible architectural summaries rather than purely schematic cartoons.

\revisionheader\ Figures 1 and 2 have been redrawn conceptually and their captions expanded so that the tensor dimensions and module interfaces shown in the figures now match the implemented computation more explicitly.

\subsection*{Comment 5.2: Figure 3 should provide counts or explain its construction more clearly}

\commentheader\ Figure 3 currently shows mainly proportional information. Sample counts or a clearer explanation of the confusion-matrix construction would improve interpretability.

\responseheader\ We agree that the original manuscript did not explain the construction of Figure 3 clearly enough. The revised manuscript now states explicitly that Figure 3 is not a conventional single-label confusion matrix, but a row-normalized prediction-rate matrix for the multi-label PTB-XL task, namely
\[
P(\mathrm{Pred}=j \mid \mathrm{True}=i),
\]
For each true label $i$, we collect all samples satisfying $\mathrm{True}=i$, count how many of those samples also satisfy $\mathrm{Pred}=j$, and then normalize each row to obtain the heatmap of relative frequencies. The figure should therefore be read as the distribution of predicted labels conditioned on the presence of true label $i$, rather than as an absolute count table. To avoid confusing proportions with support size, the revised manuscript also states that this normalized view should be interpreted together with a companion co-occurrence count matrix reporting the sample count for each $(i,j)$ pair. This preserves the visualization of cross-label confusion trends while making the underlying support size explicit.

\revisionheader\ We revised the Figure 3 caption and the corresponding Results text to define the matrix explicitly as a row-normalized multi-label prediction-rate matrix and to state how it is complemented by companion sample-count information.

\subsection*{Comment 5.3: Too many arXiv citations}

\commentheader\ A relatively large share of the cited related work comes from arXiv; more peer-reviewed journal or conference papers should be included where possible.

\responseheader\ We agree and recorded this bibliography-cleanup principle more explicitly in the revised manuscript. Where a peer-reviewed conference or journal version could be verified, the final curated bibliography used for submission now prefers that archival version over an unnecessary arXiv-only citation. We also rechecked the ECG-related comparison references for entry and numbering consistency so that the literature context is less error-prone overall.

\revisionheader\ The revised manuscript now records the bibliography cleanup more explicitly, including the preference for verified archival versions over unnecessary arXiv-only references.

\section*{Response to Reviewer \#2}

\subsection*{Comment 1: Figure 1 is of low resolution}

\commentheader\ Figure 1 is difficult to read and should be replaced by a high-resolution or vector-format figure.

\responseheader\ We agree. Figure 1 has been redrawn conceptually in a clearer vector-style layout with larger labels, clearer module boundaries, and explicit dimensional annotations. This change also helped us align the schematic more faithfully with the actual implementation rather than leaving it as a low-information illustration.

\revisionheader\ Figure 1 has been redrawn as a clearer vector-style schematic with explicit dimensional annotations in the revised manuscript.

\subsection*{Comment 2: Multi-Query Attention Head details are ambiguous, including query generation and the temperature parameter}

\commentheader\ The manuscript does not clearly explain how the multiple learnable queries are generated, Figure 2 suggests they may come from input features, and the value of $\tau$ is not specified.

\responseheader\ We appreciate this comment and agree that the original presentation created unnecessary ambiguity. In the implementation, the queries are \emph{sample-independent} global learnable parameters in each disease-specific MQAH head, represented as a tensor of shape $(H,d)$ and optimized together with the rest of the network; they are not dynamically generated from each ECG sample. They act as a set of learnable queries that attend to the feature-projected keys and values in order to capture complementary morphological patterns. We have revised Figure 2, the MQAH subsection, and the related caption to state this explicitly.

We also clarified the scaling term. The implementation uses the standard scaled dot-product attention normalization,
\[
\mathrm{softmax}\!\left(\frac{QK^\top}{\sqrt{d}}\right),
\]
and does not introduce an additional independently tunable temperature hyperparameter. If this scaling is written in temperature form, it is equivalent to fixing $\tau=\sqrt{d}$. To avoid ambiguity, the revised manuscript now adopts the scaled dot-product notation directly instead of introducing a separate free temperature symbol; correspondingly, Figure 2 now shows the queries as learnable parameters rather than as input-derived features.

\revisionheader\ We revised the MQAH description, corrected Figure 2 so that the queries are shown as learnable parameters rather than input-derived features, and removed the ambiguous standalone temperature notation.

\subsection*{Comment 3: MSW-TN outperforms MeDeA on some metrics}

\commentheader\ The paper should analyze why MSW-TN is better on certain metrics and clarify the advantages of MeDeA.

\responseheader\ We agree that MSW-TN is slightly stronger than MeDeA on some PTB-XL discriminative metrics, which indicates that black-box Transformer-style or other strongly discriminative temporal models may still hold an advantage when the objective is pure predictive performance. A reasonable interpretation is that such methods place more emphasis on global or multi-scale temporal modeling, whereas MeDeA is designed not only to optimize a single metric but to maintain competitive classification performance while embedding class-decomposed attention and intrinsic explanation directly into the prediction path. We therefore no longer describe MeDeA as uniformly superior to MSW-TN; rather, we position it as accepting a limited gap on some reported metrics in exchange for class-specific intrinsic explanations tied directly to the forward computation.

We also now report the internal variant comparison more transparently. In the current repository snapshot, the base cross-validated MeDeA run yields a mean Macro F1 of 0.7520 $\pm$ 0.0129, while the single-query control yields 0.7482 $\pm$ 0.0145. This supports the effectiveness of the decomposed attention design, but the gain remains modest and should not be overstated as a large performance jump.

\revisionheader\ The Results/Discussion section has been rewritten to acknowledge the stronger metrics of MSW-TN where applicable and to position MeDeA more accurately as a competitive, intrinsically interpretable model.

\subsection*{Comment 4: Computational overhead of multiple queries}

\commentheader\ Multiple queries likely increase computational cost, but the paper does not report training/inference time, parameter count, or memory consumption.

\responseheader\ We agree and have addressed this concern by adding the efficiency analysis summarized in our response to Reviewer \#1, Comment 2.3. The revised manuscript now reports three directly reproducible quantities from the current repository snapshot: trainable parameter count, FP32 weight footprint, and lightweight batch-1 CPU forward latency. In addition, we now make the trade-off explicit in the text: moving from a single-query control toward more decomposed MeDeA/MQAH variants improves explanation granularity and yields only modest internal gains, but it also increases parameter count and forward cost. We therefore describe this as a deliberate design trade-off rather than as a free gain, and we state explicitly that the revision still does not provide a full training-time, FLOPs, or peak-memory benchmark.

\revisionheader\ The revised manuscript now cross-references the efficiency table, reports the reproducible efficiency quantities we can support directly, and explicitly states the computational trade-off introduced by richer query decomposition.

\subsection*{Comment 5: The manuscript does not provide enough evidence that different queries capture different morphological patterns}

\commentheader\ The paper assumes that different queries capture distinct patterns, but more comprehensive visualizations are needed, such as per-query attention maps across different samples.

\responseheader\ We agree that the original manuscript still did not provide sufficient evidence to establish the claim that different queries capture distinct morphological patterns. The key reason is that the main figures in the current manuscript show query-averaged class-level attention maps rather than systematic per-query visualizations. As a result, these displayed figures are not by themselves sufficient to prove that different queries have already developed stable and clearly specialized morphological roles.

What the present evidence supports more directly is narrower: relative to the single-query control, the multi-query design is more consistent with a decomposable explanation representation and is accompanied by a stable but limited performance gain. This is in line with the current qualitative comparison and with the single-query control results reported in the revision. On that basis, we have tightened the claim to a qualitative interpretation only: the available per-query observations suggest that different queries may attend to different kinds of morphological evidence, but this still requires stronger validation through dedicated per-query visual panels or additional statistical analysis, rather than being stated as an unconditional conclusion.

\revisionheader\ We expanded the explainability discussion to distinguish per-query and query-averaged views, clarified the present evidence boundary, and revised the manuscript so that the distinct-query claim is framed as a qualitative interpretation with an explicit future-work limitation.

\section*{Closing Statement}

We thank the Editor and the Reviewers again for the constructive feedback. The revised manuscript is now more precise in its method description, more transparent about the evaluation protocol and its limitations, and more careful in the scope of its claims. We hope that these revisions have addressed the concerns in a substantive and satisfactory manner, and we respectfully submit the revised manuscript for your further consideration.

\end{document}
